cardinalR: Generating interesting high-dimensional data structures

### 2025-10-08 paper-cardinalR.Rmd
Initial codes for clustering

@@

The source material for this paper is available at [github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR).

This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template. These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], `gtools` [@gregory2023], `dplyr` [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].

>>>>

```{r, echo=TRUE}
library(fpc)

# Extract high-dimensional data
data <- five_clusts[, -5]
true_labels <- as.numeric(gsub("cluster", "", five_clusts$cluster))

# Hierarchical clustering
dist_mat <- dist(data)
hc_res <- cutree(hclust(dist_mat, method = "ward.D2"), k = 5)
```

```{r, echo=TRUE}
# K-means clustering
kmeans_res <- kmeans(data, centers = 5, nstart = 20)
```

```{r, echo=TRUE}
library(mclust)

# Model-based clustering
mclust_res <- Mclust(data, G = 5)$classification
```

```{r}
# Evaluate clustering performance
comp_kmeans <- cluster.stats(dist_mat, true_labels, kmeans_res$cluster)
comp_hclust <- cluster.stats(dist_mat, true_labels, hc_res)
comp_mclust <- cluster.stats(dist_mat, true_labels, mclust_res)

# Summarize key performance metrics
data.frame(
  Method = c("K-means", "Hierarchical", "Model-based"),
  ARI = c(comp_kmeans$corrected.rand,
          comp_hclust$corrected.rand,
          comp_mclust$corrected.rand),
  Entropy = c(comp_kmeans$entropy,
              comp_hclust$entropy,
              comp_mclust$entropy)
)
```

@@

The source material for this paper is available at [github.com/JayaniLakshika/paper-cardinalR](https://github.com/JayaniLakshika/paper-cardinalR). This article is created using \CRANpkg{knitr} [@yihui2015] and \CRANpkg{rmarkdown} [@yihui2018] in R with the `rjtools::rjournal_article` template. These `R` packages were used for this work: `cli` [@gabor2025], `tibble` [@kirill2023], `gtools` [@gregory2023], `dplyr` [@hadley2023], `stats` [@core2025], `tidyr` [@hadley2024], `purrr` [@hadley2025], `mvtnorm` [@alan2009], `geozoo` [@barret2016], and `MASS` [@venables2002].

### 2025-10-08 paper-cardinalR.tex
Initial codes for clustering

>>>>

\begin{verbatim}
library(fpc)

# Extract high-dimensional data
data <- five_clusts[, -5]
true_labels <- as.numeric(gsub("cluster", "", five_clusts$cluster))

# Hierarchical clustering
dist_mat <- dist(data)
hc_res <- cutree(hclust(dist_mat, method = "ward.D2"), k = 5)
\end{verbatim}

\begin{verbatim}
# K-means clustering
kmeans_res <- kmeans(data, centers = 5, nstart = 20)
\end{verbatim}

\begin{verbatim}
library(mclust)

# Model-based clustering
mclust_res <- Mclust(data, G = 5)$classification
\end{verbatim}

\begin{verbatim}
#>         Method ARI  Entropy
#> 1      K-means   1 1.551532
#> 2 Hierarchical   1 1.551532
#> 3  Model-based   1 1.551532
\end{verbatim}
